---
title: "DataScienceProject"
author: "Sandesh Nilas Khatiwada"
date: "2024-07-18"
output:
  html_document:
    df_print: paged
---

Cleaning the data from the file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
```

Loading data for cleaning the house data

```{r}
house2020= read_csv("pp-2020.csv", show_col_types = FALSE)

house2021 = read_csv("pp-2021.csv", show_col_types = FALSE)

house2022 = read_csv("pp-2022.csv", show_col_types = FALSE)

house2023 = read_csv("pp-2023.csv", show_col_types = FALSE)

colnames(house2020) = c("ID" , "Price", "Year", "PostCode" , "PAON", "SAON", "FL", "House Num", "Flat", "Street Name",
                     
                     "Locality", "Town" , "District", "County", "Type1", "Type2" )

colnames(house2021) = c("ID" , "Price", "Year", "PostCode" , "PAON", "SAON", "FL", "House Num", "Flat", "Street Name",
                     
                     "Locality", "Town" , "District", "County", "Type1", "Type2")

colnames(house2022) = c("ID" , "Price", "Year", "PostCode" , "PAON", "SAON", "FL", "House Num", "Flat", "Street Name",
                     
                     "Locality", "Town" , "District", "County" , "Type1", "Type2")

colnames(house2023) = c("ID" , "Price", "Year", "PostCode" , "PAON", "SAON", "FL", "House Num", "Flat", "Street Name",
                     
                     "Locality", "Town" , "District", "County" , "Type1", "Type2")

```

Cleaning the data such that it will omit the null value and then only saving the unique data

```{r}

housingPrice = rbind(house2020,house2021,house2022,house2023) %>%
  na.omit() %>%
  distinct() %>%
  as_tibble()

```

creating the cleaned csv file where there will be no null value in house price data

```{r}
write_csv(housingPrice, "cleaned_price_data.csv")
```

filter all the data for Bristol and Cornwall

```{r}
selectedCounty <- housingPrice %>% 
  filter(County %in% c('CITY OF BRISTOL','CORNWALL'))
print(selectedCounty)
```

now grouping the dataset according to year 

```{r}
selectedCounty$Year <- as.Date(selectedCounty$Year)

# extracting year from the date 
selectedCounty$y <- year(selectedCounty$Year)

# grouping data acording to year 
grouped_housing <- selectedCounty %>% 
  group_by(y,County) %>%
  summarise(avg_price = mean(Price))

print(grouped_housing)



```

now I am creating the visualization for housing

```{r}
#line chart for year 2020-2023 for average of both county 
ggplot(grouped_housing, aes(x=y, y= avg_price, colour = County, group = County) ) +
  geom_line()+
  geom_point()+
  labs(title = "Property Prices Over Time by County (2020-23)",
       x = "Year",
       y = "Average Price")+theme_minimal()

```

now creating the data for only year 2022 

```{r}

# Filter data for the year 2022
data_2022 <- selectedCounty %>%
  filter(y == 2022)

#box plot 
ggplot(data = data_2022, aes(x = County, y = Price, color = County)) +
  geom_boxplot() +
  labs(title = "Property Prices in 2022 by City",
       x = "County",
       y = "Price") +
  theme_minimal()


```


```{r}
#barChart 
ggplot(data = data_2022, aes(x = County, y = Price, fill = County)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Property Prices in 2022 by City",
       x = "County",
       y = "Total Price") +
  theme_minimal()


```








